---
title: "p8105_hw6_rj2543"
author: "rj2543"
date: "November 19, 2018"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

#library(HH)
#library("leaps")
#library(purrr)

set.seed(1)
```

# Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository. You can read their accompanying article.

```{r data import}
homicide = read_csv("./data/homicide-data.csv") %>% 
  janitor::clean_names()

homicide
```

Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. Modifiy victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r data wrangling}
homi = homicide %>% 
  unite(city_state, city, state, sep = ", ") %>% 
  mutate(solved = ifelse(disposition == "Open/No arrest", 0, 1)) %>% # 1: Closed without arrest/Closed by arrest; 0: Open/No arrest
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>% 
  mutate(victim_race = ifelse(victim_race == "White", "white", "non-white")) %>% # modify 'victim_race' to have categories
  mutate(victim_race = forcats::fct_relevel(victim_race, ref = "white")) %>% # with white as the reference category
  mutate(victim_age = suppressWarnings(as.numeric(victim_age)))

homi
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r Baltimore}
baltimore = homi %>% 
  filter(city_state == "Baltimore, MD")

reg_baltimore = glm(solved~victim_age+victim_sex+victim_race, family = binomial(link = "logit"), data = baltimore)

OR_CI = rbind(exp(confint(reg_baltimore, "victim_racenon-white")))

reg_baltimore %>%  
  broom::tidy() %>% 
  filter(term == "victim_racenon-white") %>% 
  mutate(OR_estimate = exp(estimate), OR_CI_low = OR_CI[1], OR_CI_up = OR_CI[2]) %>%
  select(term, logOR_estimate = estimate, OR_estimate, OR_CI_low, OR_CI_up, p.value)

#baltimore %>% 
  #modelr::add_predictions(reg_baltimore) %>% 
  #mutate(fitted_prob = boot::inv.logit(pred))
```


Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r every city, message = FALSE, warning = FALSE}
or_ci = function(city){
  city_info = homi %>% 
    filter(city_state == city)
  
  reg_city = glm(solved~victim_age+victim_sex+victim_race, family = binomial(link = "logit"), data = city_info)
  
  tidy_city = reg_city %>% 
    broom::tidy() %>% 
    filter(term == "victim_racenon-white")
  
  OR_estimate = exp(tidy_city$estimate)
  OR_CI = rbind(exp(confint(reg_city, "victim_racenon-white")))
  
  return(tibble(
    OR_estimate,
    OR_CI_low = OR_CI[1],
    OR_CI_up = OR_CI[2]
  ))
}

OR_CI_city = homi %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(info = map(city_state, or_ci)) %>% 
  select(city_state, info) %>% 
  unnest()

OR_CI_city %>% 
  knitr::kable(digits = 5)
  
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r plot}
OR_CI_city %>% 
  mutate(city_state = fct_reorder(city_state, OR_estimate)) %>% 
  ggplot(aes(x = city_state, y = OR_estimate)) +
  geom_point() +
  geom_errorbar(aes(x = city_state, ymin = OR_CI_low, ymax = OR_CI_up), width = 0.2, alpha = 0.5) + # add error bars based on upper and lower limits of CI for each city
  labs(
    title = "Estimates and CIs for each city",
    x = "City",
    y = "Estimate and CI"
  ) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10)) # make city_state names readable
```



# Problem 2

In this probelm, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

* babysex: baby’s sex (male = 1, female = 2)

* bhead: baby’s head circumference at birth (centimeters)

* blength: baby’s length at birth (centimeteres)

* bwt: baby’s birth weight (grams)

* delwt: mother’s weight at delivery (pounds)

* fincome: family monthly income (in hundreds, rounded)

* frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)

* gaweeks: gestational age in weeks

* malform: presence of malformations that could affect weight (0 = absent, 1 = present)

* menarche: mother’s age at menarche (years)

* mheigth: mother’s height (inches)

* momage: mother’s age at delivery (years)

* mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)

* parity: number of live births prior to this pregnancy

* pnumlbw: previous number of low birth weight babies

* pnumgsa: number of prior small for gestational age babies

* ppbmi: mother’s pre-pregnancy BMI

* ppwt: mother’s pre-pregnancy weight (pounds)

* smoken: average number of cigarettes smoked per day during pregnancy

* wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r data import and clean}
birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = as.factor(babysex), frace = as.factor(frace), malform = as.factor(malform), mrace = as.factor(mrace))

birthweight
```


Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r regression}
birthweight %>% 
  ggplot(aes(x = bwt)) +
  geom_histogram()

reg = lm(bwt~., data = birthweight)

summary(reg)

reg_stepwise = step(reg, direction = "both")

summary(reg_stepwise)

reg_bwt = lm(bwt~babysex+bhead+blength+delwt+gaweeks+mheight+mrace+ppwt+smoken, data = birthweight)

summary(reg_bwt)

birthweight %>% 
  modelr::add_residuals(., reg_bwt) %>% 
  modelr::add_predictions(., reg_bwt) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5, size = 1) +
  geom_abline(aes(slope = 0, intercept = 0), color = "blue")

```


Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)

```{r blength and gaweeks}
reg1 = lm(bwt~blength+gaweeks, data = birthweight)

summary(reg1)

reg1 %>% 
  broom::tidy()

cv_df = crossv_mc(birthweight, 100) %>% 
  mutate(reg_mod = map(train, ~lm(bwt~babysex+bhead+blength+delwt+gaweeks+mheight+mrace+ppwt+smoken, data = .x)),
         reg1_mod = map(train, ~lm(bwt~blength+gaweeks, data = .x))) %>% 
  mutate(rmse_reg = map2_dbl(reg_mod, test, ~rmse(model = .x, data = .y)),
         rmse_reg1 = map2_dbl(reg1_mod, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(rmse_reg, rmse_reg1) %>%  
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```


* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these.

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
reg2 = lm(bwt~bhead*blength*babysex, data = birthweight)

summary(reg2)

reg2 %>% 
  broom::tidy()

cv_df = crossv_mc(birthweight, 100) %>% 
  mutate(reg_mod = map(train, ~lm(bwt~babysex+bhead+blength+delwt+gaweeks+mheight+mrace+ppwt+smoken, data = .x)),
         reg2_mod = map(train, ~lm(bwt~bhead*blength*babysex, data = .x))) %>% 
  mutate(rmse_reg = map2_dbl(reg_mod, test, ~rmse(model = .x, data = .y)),
         rmse_reg2 = map2_dbl(reg2_mod, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(rmse_reg, rmse_reg2) %>%  
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```


Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

